---
title: "Time Series Model Generation and ACVF/ACF Computation Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time Series Model Generation and ACVF/ACF Computation Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=7, fig.height=6) 
```

```{r setup}
library(TsimSeriesACF)
```

This vignette serves as an example on how to use the functions within the `TsimSeriesACF` package for simulating autoregressive (AR), moving-average (MA), and autoregressive moving-average (ARMA) models, compute their corresponding theoretical and sample autocovariance (ACVF) and autocorrelation (ACF) function, and helper functions to do so:

-   Model Generation `TsimSeriesACF::gen_arma()`,

-   Converting ARMA models into infinite order MA models `TsimSeriesACF::arma_to_ma()`,

-   Determining causality of a model `TsimSeriesACF::is_causal()`, 

-   Computing theoretical ACVF and ACF of a model `TsimSeriesACF::get_theoretical_acvf()` and `TsimSeriesACF::get_theoretical_acf()` respectively.

-   Computing sample ACVF and ACF of a model `TsimSeriesACF::get_sample_acvf()` and `TsimSeriesACF::get_sample_acf()` respectively.

The following examples will be sampling time series data from the following ARMA(1, 2), AR(1), and MA(2) models:
\[X_t = 0.5X_{t - 1} + w_t + 0.5w_{t - 1} + 0.25w_{t - 2},\]
\[X_t = 0.5X_{t - 1} + w_t, \text{ and}\]
\[X_t = w_t + 0.5w_{t - 1} + 0.25w_{t - 2},\]

where $w_t \overset{\text{i.i.d.}}{\sim} N(\mu = 0, \sigma_W^2 = 1).$ 

Note, while we will generate data and compare results for all three models, we omit examples of utilizing the helper functions `TsimSeriesACF::arma_to_ma()` and `TsimSeriesACF::is_causal()` for AR and MA models as they are similarly utilized.

## Generating Time Series Data for all models and plotting them

```{r}
set.seed(1)
n <- 50
phi <- 0.5
theta <- c(0.5, 0.25)
wn <- rnorm(n)

ar <- gen_arma(n, wn, ar_coefs = phi, ma_coefs = NULL)
ma <- gen_arma(n, wn, ar_coefs = NULL, ma_coefs = theta)
arma <- gen_arma(n, wn, ar_coefs = phi, ma_coefs = theta)

plot.ts(ar$data, main = "AR(1) Simulated Data", ylab = "X_t", xlab = "t")
plot.ts(ma$data, main = "MA(2) Simulated Data", ylab = "X_t", xlab = "t")
plot.ts(arma$data, main = "ARMA(1,2) Simulated Data", ylab = "X_t", xlab = "t")
```

## Converting ARMA Model to Infinite-order MA Model

```{r}
converted_arma <- arma_to_ma(arma, max_lag = 10)
```

## Checking if Model is Causal

```{r}
is_causal(arma)
```

## Calculating Theoretical and Sample ACVFs and ACFs

```{r}
# For ARMA
arma_t_acvf <- get_theoretical_acvf(arma, sigma = 1, max_lag = 20)
arma_t_acf <- get_theoretical_acf(arma, sigma = 1, max_lag = 20)
arma_s_acvf <- get_sample_acvf(arma, max_lag = 20)
arma_s_acf <- get_sample_acf(arma, max_lag = 20)

# For AR
ar_t_acvf <- get_theoretical_acvf(ar, sigma = 1, max_lag = 20)
ar_t_acf <- get_theoretical_acf(ar, sigma = 1, max_lag = 20)
ar_s_acvf <- get_sample_acvf(ar, max_lag = 20)
ar_s_acf <- get_sample_acf(ar, max_lag = 20)

# For MA
ma_t_acvf <- get_theoretical_acvf(ma, sigma = 1, max_lag = 20)
ma_t_acf <- get_theoretical_acf(ma, sigma = 1, max_lag = 20)
ma_s_acvf <- get_sample_acvf(ma, max_lag = 20)
ma_s_acf <- get_sample_acf(ma, max_lag = 20)
```

## Plotting Theoretical ACVF and ACF

Here we visualize the sample and theoretical ACF and ACVF for each model.

```{r}
# function to compare plots
plot_compare <- function(theoretical, sample, main, ylab) {
  lags <- 0:(length(theoretical) - 1)
  plot(lags, theoretical, type = "b", col = "blue", pch = 19,
       main = main, xlab = "Lag", ylab = ylab,
       ylim = range(c(theoretical, sample), na.rm = TRUE))
  lines(lags, sample, type = "b", col = "red", pch = 17)
  abline(h = 0)
  legend("topright", legend = c("Theoretical", "Sample"),
         col = c("blue", "red"), pch = c(19, 17), lty = 1, bty = "n")
}

# ARMA(1,2) plots
plot_compare(arma_t_acvf, arma_s_acvf,
             main = "ARMA(1,2) ACVF: Theoretical vs Sample",
             ylab = "ACVF")

plot_compare(arma_t_acf, arma_s_acf,
             main = "ARMA(1,2) ACF: Theoretical vs Sample",
             ylab = "ACF")

# AR(1) plots
plot_compare(ar_t_acvf, ar_s_acvf,
             main = "AR(1) ACVF: Theoretical vs Sample",
             ylab = "ACVF")

plot_compare(ar_t_acf, ar_s_acf,
             main = "AR(1) ACF: Theoretical vs Sample",
             ylab = "ACF")

# MA(2) plots
plot_compare(ma_t_acvf, ma_s_acvf,
             main = "MA(2) ACVF: Theoretical vs Sample",
             ylab = "ACVF")

plot_compare(ma_t_acf, ma_s_acf,
             main = "MA(2) ACF: Theoretical vs Sample",
             ylab = "ACF")
```

## Model Comparisons

We can see the following defining characteristics among the example models:

- The theoretical MA(2) ACVF and ACF go to zero after lag 2, as expected for MA(q) processes.

- The theoretical AR(1) and ARMA(1, 2) ACVF and ACF converge to zero as lag goes to infinity, also expected for AR(p) and ARMA(p, q) processes.

- Sample ACVF and ACF follow their theoretical counterparts, though offset, and exhibit sinusoidal behavior for large lags, reflecting the noise found in samples.

## Generative AI Declaration

Generative AI was used to assist in the writing, testing, and documentation of some functions and test cases.

For the package compilation phase, generative AI was consulted for conventions
on package compilation, and to scan for typos.

# Bibliography


