---
title: "Time Series Model Generation and ACVF/ACF Computation Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=7, fig.height=6) 
```

```{r setup}
library(sta380)
```

This vignette serves as an example on how to use the functions within the `sta380` package for simulating autoregressive (AR), moving-average (MA), and autoregressive moving-average (ARMA) models, compute their corresponding theoretical and sample autocovariance (ACVF) and autocorrelation (ACF) function, plot their data, and related functions:

-   Model Generation `sta380::gen_arma()`,

-   Converting ARMA models into infinite order MA models `sta380::arma_to_ma()`,

-   Determining causality of a model `sta380::is_causal()`, 

-   Computing theoretical ACVF and ACF of a model `sta380::get_theoretical_acvf()` and `sta380::get_theoretical_acf()` respectively.

-   Computing sample ACVF and ACF of a model...

The following examples will be sampling time series data from the following ARMA(1, 2), AR(1), and MA(2) models:
\[X_t = 0.5X_{t - 1} + w_t + 0.5w_{t - 1} + 0.25w_{t - 2},\]
\[X_t = 0.5X_{t - 1} + w_t, \text{ and}\]
\[X_t = w_t + 0.5w_{t - 1} + 0.25w_{t - 2},\]

where $w_t \overset{\text{i.i.d.}}{\sim} N(\mu = 0, \sigma_W^2 = 1).$ 

Note, while we will generate data and compare results for all three models, we omit examples of utilizing the helper functions `sta380::arma_to_ma()` and `sta380::is_causal()` for AR and MA models as they are similarly utilized.

# 

## Generating Time Series Data for all models

```{r}
n <- 10
phi <- 0.5
theta <- c(0.5, 0.25)
wn <- rnorm(n)

ar <- gen_arma(n, wn, ar_coefs = phi, ma_coefs = NULL)
ma <- gen_arma(n, wn, ar_coefs = NULL, ma_coefs = theta)
arma <- gen_arma(n, wn, ar_coefs = phi, ma_coefs = theta)

# Sampled data from model
plot.ts(arma$data)

# Sampled white noise used in model
arma$wn
```

## Converting ARMA Model to Infinite-order MA Model

```{r}
converted_arma <- arma_to_ma(arma, max_lag = 10)
```

## Checking if Model is Causal

```{r}
is_causal(arma)
```

## Calculating Theoretical and Sample ACVFs and ACFs

```{r}
# For ARMA
arma_t_acvf <- get_theoretical_acvf(arma, sigma = 1, max_lag = 10)
arma_t_acf <- get_theoretical_acf(arma, sigma = 1, max_lag = 10)
# arma_s_acvf <- get_sample_acvf(arma, max_lag = 10)
# arma_s_acf <- get_sample_acf(arma, max_lag = 10)

# For AR
ar_t_acvf <- get_theoretical_acvf(ar, sigma = 1, max_lag = 10)
ar_t_acf <- get_theoretical_acf(ar, sigma = 1, max_lag = 10)
# ar_s_acvf <- get_sample_acvf(ar, max_lag = 10)
# ar_s_acf <- get_sample_acf(ar, max_lag = 10)

# For MA
ma_t_acvf <- get_theoretical_acvf(ma, sigma = 1, max_lag = 10)
ma_t_acf <- get_theoretical_acf(ma, sigma = 1, max_lag = 10)
# ma_s_acvf <- get_sample_acvf(ma, max_lag = 10)
# ma_s_acf <- get_sample_acf(ma, max_lag = 10)
```

## Plotting Theoretical ACVF and ACF

# Comparison of All Three

Looking at the results from all three samples, it appears that the Gibbs sampler gives the closest result to what we tried to simulate (although this has not been statistically verified.)

```{r}
sims <- list(sim_attempt1, sim_attempt2, sim_attempt3)

# Means
sapply(sims, function(m) mean(m[,1]))

# Standard Deviations
sapply(sims, function(m) sd(m[,1]))

# Correlations
sapply(sims, function(m) cor(m[,1], m[,2]))
```

# Bibliography


